---
title: "Process and merge data"
author: "Suparna Chaudhry and Andrew Heiss"
date: "Last run: `r format(Sys.time(), '%F')`"
output: 
  html_document:
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knit_print.data.frame <- function(x, ...) {
  res <- paste(c('', '', kable_styling(kable(x, booktabs = TRUE))), collapse = '\n')
  asis_output(res)
}

registerS3method("knit_print", "data.frame", knit_print.data.frame)
registerS3method("knit_print", "grouped_df", knit_print.data.frame)

knitr::opts_chunk$set(fig.retina = 3,
                      tidy.opts = list(width.cutoff = 120),  # For code
                      options(width = 90),  # For output
                      fig.asp = 0.618, fig.width = 7, 
                      fig.align = "center", out.width = "85%")

options(dplyr.summarise.inform = FALSE,
        knitr.kable.NA = "")
```

```{r load-libraries-data, message=FALSE, warning=FALSE}
library(tidyverse)
library(targets)
library(scales)
library(patchwork)
library(DT)
library(naniar)
library(here)

# Generated via random.org
set.seed(9936)

# Load data
# Need to use this withr thing because tar_read() and tar_load() need to see the
# _targets folder in the current directory, but this .Rmd file is in a subfolder
withr::with_dir(here::here(), {
  source(tar_read(plot_funs))
  
  tar_load(skeleton)
  tar_load(chaudhry_clean)
  tar_load(pts_clean)
  tar_load(latent_hr_clean)
  tar_load(killings_all)
  tar_load(vdem_clean)
  tar_load(wdi_clean)
  tar_load(un_pop)
  tar_load(un_gdp)
  tar_load(ucdp_prio_clean)
  
  tar_load(panel)
})
```

# Data processing

## Country-year skeleton

We use [Gleditsch-Ward country codes](http://ksgleditsch.com/data-4.html) to identify each country across the different datasets we merge. We omit a bunch of things though:

- We omit microstates
- Because the World Bank doesn't include it in the WDI, we omit Taiwan (713)
- We only use countries in Suparna's anti-NGO law data

To get consistency in country codes, we do this:

- When converting GW codes to COW codes, [following Gleditsch and Ward](https://www.andybeger.com/states/articles/differences-gw-cow.html), we treat post-2006 Serbia as 345 (a continuation of Serbia & Montenegro). And we also treat Serbia as a continuation of Yugoslavia with 345 (following V-Dem, which does that too)
- In both COW and GW codes, modern Vietnam is 816, but `countrycode()` thinks the COW code is 817, which is old South Vietnam (see [issue](https://github.com/vincentarelbundock/countrycode/issues/16)), so we use `custom_match` to force 816 to recode to 816.
- Also, following V-Dem, we treat Czechoslovakia (GW/COW 315) and Czech Republic (GW/COW 316) as the same continuous country (V-Dem has both use ID 157).

We have `r panel %>% distinct(gwcode) %>% nrow()` countries in this data, spanning `r panel %>% distinct(year) %>% nrow()` possible years. Here's a lookup table of all the countries included:

```{r show-final-countries}
skeleton$skeleton_lookup %>% 
  filter(gwcode %in% unique(chaudhry_clean$gwcode)) %>% 
  select(-years_included) %>% 
  datatable()
```

## Chaudhry NGO restrictions

We create several indexes for each of the categories of regulation, following Christensen and Weinstein’s classification:

- `entry` (Q2b, Q2c, Q2d; 3 points maximum, actual max = 3 points maximum): barriers to entry
  - Q2c is reversed, so not being allowed to appeal registration status earns 1 point.
  - Q2a is omitted because it’s benign
- `funding` (Q3b, Q3c, Q3d, Q3e, Q3f; 5 points maximum, actual max = 4.5): barriers to funding
  - Q3a is omitted because it’s benign
  - Scores that range between 0–2 are rescaled to 0–1 (so 1 becomes 0.5)
- `advocacy` (Q4a, Q4c; 2 points maximum, actual max = 2): barriers to advocacy
  - Q4b is omitted because it’s not a law
  - Scores that range between 0–2 are rescaled to 0–1 (so 1 becomes 0.5)
- `barriers_total` (10 points maximum, actual max = 8.5): sum of all three indexes

These indexes are also standardized by dividing by the maximum, yielding the following variables:

- `entry_std`: 1 point maximum, actual max = 1
- `funding_std`: 1 point maximum, actual max = 1
- `advocacy_std`: 1 point maximum, actual max = 1
- `barriers_total_std`: 3 points maximum, actual max = 2.5

```{r glimpse-chaudhry-ngos}
glimpse(chaudhry_clean)
```

## Political Terror Scale (PTS)

We use data from the [Political Terror Scale (PTS) project](http://www.politicalterrorscale.org/) to measure state repression. This project uses reports from the US State Department, Amnesty International, and Human Rights Watch and codes political repression on a scale of 1-5:

- **Level 1**: Countries under a secure rule of law, people are not imprisoned for their view, and torture is rare or exceptional. Political murders are extremely rare.
- **Level 2**: There is a limited amount of imprisonment for nonviolent political activity. However, few persons are affected, torture and beatings are exceptional. Political murder is rare.
- **Level 3**: There is extensive political imprisonment, or a recent history of such imprisonment. Execution or other political murders and brutality may be common. Unlimited detention, with or without a trial, for political views is accepted.
- **Level 4**: Civil and political rights violations have expanded to large numbers of the population. Murders, disappearances, and torture are a common part of life. In spite of its generality, on this level terror affects primarily those who interest themselves in politics or ideas.
- **Level 5**: The terrors of Level 4 have been extended to the whole population. The leaders of these societies place no limits on the means or thoroughness with which they pursue personal or ideological goals.

Following Gohdes and Carey, we use the State Department score, unless it's missing, in which case we use Amnesty's score.

```{r glimpse-pts}
glimpse(pts_clean)
```

## Latent Human Rights Protection Scores

We also use Chris Fariss's [Latent Human Rights Protection Scores](https://doi.org/10.7910/DVN/RQ85GK), which are estimates from fancy Bayesian models that capture a country's respect for physical integrity rights. We use the posterior mean of the latent variable from the model (θ), and we keep the standard deviation of θ just in case we want to fancier things in the future and incorporate the uncertainty of θ in the models. We rename θ to `latent_hr_mean`.

In this measure, high values represent strong respect for physical integrity rights:

```{r showh-lhr-min-max}
latent_hr_clean %>% 
  filter(latent_hr_mean == max(latent_hr_mean) | 
           latent_hr_mean == min(latent_hr_mean)) %>% 
  left_join(select(skeleton$skeleton_lookup, gwcode, country),
            by = "gwcode") %>% 
  select(country, everything())
```

```{r glimpse-lhr}
glimpse(latent_hr_clean)
```

## Gohdes and Carey journalist deaths

We use data from [Gohdes and Carey (2017)](https://doi.org/10.1177/0022343316680859) for the number of journalists murdered between 2002 and 2014. They provide two `.Rdata` files in their replication materials—one is a list of all journalist deaths, and one is a TSCS dataset they use in their main analysis, with summary values of journalist deaths for each country/year. Their panel data does not include 2002 for some reason, so we ignore the panel data and recreate the summary values by hand here.

```{r glimpse-gh-journalists}
glimpse(killings_all)
```

## Varieties of Democracy (V-Dem)

We use a bunch of variables from the [Varieties of Democracy project](https://www.v-dem.net/en/):

- **Civil society stuff**
  - CSO entry and exit: `v2cseeorgs`
  - CSO repression: `v2csreprss`
  - CSO consultation: `v2cscnsult`
  - CSO participatory environment: `v2csprtcpt`
  - CSO women's participation: `v2csgender`
  - CSO anti-system movements: `v2csantimv`
  - Core civil society index (entry/exit, repression, participatory env): `v2xcs_ccsi`
- **Freedom of expression stuff**
  - Freedom of expression index: `v2x_freexp`
  - Freedom of expression and alternative sources of information index: `v2x_freexp_altinf`
  - Freedom of academic and cultural expression: `v2clacfree`
  - Media self-censorship: `v2meslfcen`
- **Repression stuff**
  - Religious organization repression: `v2csrlgrep`
  - Govt censorship effort - media: `v2mecenefm`
  - Harassment of journalists: `v2meharjrn`
- **Rights indexes**
  - Civil liberties index: `v2x_civlib`
  - Physical violence index: `v2x_clphy`
  - Private civil liberties index: `v2x_clpriv`
  - Political civil liberties index: `v2x_clpol`
- **Democracy and governance stuff**
  - Polyarchy index (for electoral democracies): `v2x_polyarchy`
  - Liberal democracy index (for democracies in general): `v2x_libdem`
  - Regimes of the world category: `v2x_regime`
  - Political corruption index: `v2x_corr` (less to more, 0-1) (public sector + executive + legislative + judicial corruption)
  - Rule of law index: `v2x_rule`


```{r glimpse-vdem}
glimpse(vdem_clean)
```

## World Bank development indicators

We don't really use anything from the World Bank's data except for population data for Kosovo.

```{r wlimpse-wdi}
glimpse(wdi_clean)
```

## UNData

The reason we don't just use WDI data for GDP and % of GDP from trade is that the WDI data is incomplete (especially pre-1990, but that's not an issue in this project) To get around that, we create our own GDP and trade measures using data directly from the UN (at [UNData](https://data.un.org/)). They don't have a neat API like the World Bank, so you have to go to their website and export the data manually.

We collect three variables: [GDP at constant 2015 prices](http://data.un.org/Data.aspx?q=gdp&d=SNAAMA&f=grID%3a102%3bcurrID%3aUSD%3bpcFlag%3a0), [GDP at current prices](http://data.un.org/Data.aspx?q=gdp&d=SNAAMA&f=grID%3a101%3bcurrID%3aUSD%3bpcFlag%3a0), and [population](https://population.un.org/wpp/Download/Standard/Population/).

```{r glimpse-un-pop}
glimpse(un_pop)
```

```{r glimpse-un-gdp}
glimpse(un_gdp)
```

## UCDP/PRIO Armed Conflict

Following [Gohdes and Carey (2017)](https://doi.org/10.1177/0022343316680859), we use [UCDP/PRIO Armed Conflict data](https://ucdp.uu.se/downloads/index.html#armedconflict) to create an indicator marking if a country-year was involved in armed conflict that resulted in at least 25 battle-related deaths.

```{r glimpse-ucdp-prio}
glimpse(ucdp_prio_clean)
```

## Final clean combined data

Finally, we join everything together in one nice clean panel:

```{r glimpse-combined}
glimpse(panel)
```

\ 


# Basic data details

```{r show-country-years-summary}
panel %>% 
  summarize(`Number of countries` = length(unique(gwcode)),
            `Number of years` = length(unique(year)),
            `First year` = min(year),
            `Last year` = max(year)) %>% 
  kbl(align = "c") %>% 
  kable_styling()
```

Countries across years:

```{r show-country-year-count}
panel %>% 
  count(Year = year, name = "Number of countries") %>% 
  kbl(align = "c") %>% 
  kable_styling()
```

\ 


# Missingness

We were super careful to ensure that our data is as complete as possible.

```{r plot-missing}
panel %>% 
  vis_miss() +
  theme_ngo() +
  theme(axis.text.x = element_text(hjust = 0, angle = 45),
        panel.grid.major = element_blank())
```

There are only a tiny handful of rows with missing data: Serbia pre-2006.

```{r show-missing-rows, class.source="fold-show"}
panel %>% 
  filter(is.na(advocacy)) %>% 
  select(gwcode, year, country, iso3, advocacy, entry, funding)
```

That's because our anti-NGO data doesn't include data for those years. That's fine, though. In models where we look at anti-NGO laws, we exclude those rows and only include rows where our `laws` indicator variable is TRUE. In models where we look at the general civil society environment, we do include these rows because V-Dem has data for them.

There are just two instances of missing data with PTS scores:

```{r show-missing-pts, class.source="fold-show"}
panel %>% 
  filter(is.na(PTS)) %>% 
  select(gwcode, year, country, iso3, PTS, PTS_factor)
```

We just ignore them, I guess.

With that, here's our actual final cleaned data:

```{r plot-missing-laws}
panel %>% 
  filter(laws) %>% 
  vis_miss() +
  theme_ngo() +
  theme(axis.text.x = element_text(hjust = 0, angle = 45),
        panel.grid.major = element_blank())
```

Beautiful.

\ 


# `targets` pipeline

Here's the general process for building and running this analysis. This is all done with [the magical **`targets`** package](https://docs.ropensci.org/targets/), which orchestrates all the dependencies automatically.

```{r show-targets-pipeline, echo=FALSE}
withr::with_dir(here::here(), {
  targets::tar_glimpse()
})
```


\ 


# Actual code

All the data processing is handled with dataset-specific functions that live in `R/funs_data-cleaning.R`, which **`targets`** then runs as needed. For the sake of transparency, here's that code:

```{r, code=xfun::read_utf8(here::here("R", "funs_data-cleaning.R")), eval=FALSE, class.source="fold-show"}
```
